### 2020-09-09 1.P259.pDC_data.cleaning.Rmd
init public manuscript repo

---
title: 'P259: Combined data cleaning'
subtitle: "Dendritic cells (pDC)"
author: "Kim Dill-McFarland, kadm@uw.edu"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
always_allow_html: true
---
# Background

P259.1: pDC were isolated from 4 healthy donors' whole blood or buffy coat. Cells were adapted to media with IL3 and then cultured with eosinophil (EOS) supernatant and/or human rhinovirus (RV).

P259.2: pDC were isolated from 7 asthmatic donors' whole blood, 4 of which were on Anti-IL5 therapy. Cells were adapted to media with IL3 and then cultured with or without RV.

The purpose of this workflow is to complete quality control and data cleaning of metadata and RNA-seq libraries generated from the above experiments. This includes 1) removing low coverage libraries, 2) filtering rare genes, 3) removing outlying libraries, and 4) normalizing for RNA composition.

# Setup
Load packages

```{r message=FALSE, warning=FALSE}
# Data manipulation and figures
library(tidyverse)
  #Multi-panel figures
  library(cowplot)
  # Modify ggplot figures to non-overlapping text labels
  library(ggrepel)
  # Modify ggplot data order within facets
  library(drlib)
  # Plot log scales
  library(scales)
# Reading Excel files
library(readxl)
# Working with dates
library(lubridate)

# Empirical analysis of digital gene expression data
## Data normalization
library(edgeR)

# Print pretty table to knit file
library(knitr)
library(kableExtra)
  options(knitr.kable.NA = '')

#Define colors for plots
group.cols <- c("none:none"="#dadaeb",
               "none:AntiIL5"="#9e9ac8",
               "none:EOS.supp"="#54278f",
               "HRV:none"="#c7e9c0",
               "HRV:AntiIL5"="#74c476",
               "HRV:EOS.supp"="#006d2c",
               "flu:none"="#fdae6b",
               "flu:AntiIL5"="#e6550d")
samp.cols <- c("AC1"="#969696",
               "AC2"="#a6cee3",
               "AC3"="#1f78b4",
               "AC4"="#b2df8a",
               "AC5"="#33a02c",
               "AT1"="#fb9a99",
               "AT2"="#e31a1c",               
               "AT3"="#fdbf6f",
               "AT4"="#ff7f00",
               "donor1"="#cab2d6",
               "donor2"="#6a3d9a",
               "donor3"="#ffff99",
               "donor4"="#b15928")
```

Set seed

```{r}
set.seed(4389)
```

# Read in and format data 
## Metadata

```{r echo=FALSE, message=FALSE}
meta <- read_csv("data_clean/P259_pDC_metadata.csv") %>% 
  #Set factosr and orders
  mutate(IL5 = factor(IL5, levels = c("none","AntiIL5","EOS.supp")),
         virus = factor(virus, levels=c("none","HRV")),
         virus.detail = factor(virus.detail,
                               levels=c("none","oldHRV","newHRV"))) %>% 
  arrange(libID)
```

## Counts table

Read in counts table from P259.1 and P259.2 (`data_raw/*combined_counts.csv`) and then combine to one data frame.

```{r echo=FALSE, message=FALSE}
count.raw1 <- read_csv("data_raw/P216-1_CCLKLANXX_180806_combined_counts.csv") %>% 
  # Remove '_CE03HANXX' suffix from variable names
  rename_at(.vars = vars(ends_with("_CCLKLANXX")),
            .funs = ~sub("_CCLKLANXX", "", .))
```

```{r echo=FALSE, message=FALSE}
count.raw2 <- read_csv("data_raw/P259-2_CE03HANXX_191107_combined_counts.csv") %>% 
  # Remove '_CE03HANXX' suffix from variable names
  rename_at(.vars = vars(ends_with("_CE03HANXX")),
            .funs = ~sub("_CE03HANXX", "", .))
```

```{r echo=FALSE, message=FALSE}
count.all <- full_join(count.raw1, count.raw2,
                       by = "geneName")
```

Read in and filter associated gene key (`data_raw/2020.06.18_HGNC.gene.key.txt`, downloaded from <https://www.genenames.org/download/custom/>) to protein coding (pc) genes that occur in the count data set and have valid hgnc symbols.

```{r echo=FALSE, message=FALSE}
key <- read_tsv("data_raw/2020.06.18_HGNC.gene.key.txt",
                na=c("")) %>% 
  # rename variable to match count data
  dplyr::rename(geneName = "Ensembl gene ID",
                hgnc_symbol = "Approved symbol",
                gene_biotype = "Locus group") %>% 
  #remove genes w/o ensembl ID
  drop_na(geneName) %>% 
  #collapse multi-annotations
  group_by(geneName) %>% 
  mutate(across(c(gene_biotype, hgnc_symbol), 
                ~paste(unique(.), collapse=", "))) %>% 
  distinct()

key.pc <- key %>% 
  # Keep protein coding genes only
  filter(gene_biotype == "protein-coding gene") %>% 
  # Keep genes in counts data
  filter(geneName %in% count.all$geneName) %>% 
  # Remove duplicate entries
  distinct(geneName, .keep_all=TRUE) %>% 
  arrange(geneName)
```

Then filter the count data to pc genes and save.

```{r echo=FALSE, message=FALSE}
count.pDC <- count.all %>% 
  filter(geneName %in% key.pc$geneName) %>% 
  arrange(geneName) %>% 
  #Order columns and keep only pDC data
  select(geneName, all_of(meta$libID))
```

```{r}
#Save counts table
write_csv(count.pDC, "data_clean/P259_pDC_counts.csv")
```

```{r echo=FALSE}
#Remove intermed files
rm(count.raw1, count.raw2, count.all, key)
```

#### Verify data order

To ensure that all data match, check that all libraries in the count data are present and in the same order in the metadata. And vice versa.

```{r echo=FALSE, message=FALSE}
identical(meta$libID, colnames(count.pDC)[-1])
```

## Sample summary

At this point, the following objects exist in our environment. Protein coding = pc

```{r echo=FALSE}
ls()
```

And the sample numbers are

```{r echo=FALSE}
meta %>% 
  group_by(experiment, asthma,IL5,virus) %>% 
  tally() %>% 
  arrange(experiment, asthma,IL5,virus) %>%

  kable(align="c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>% 
  collapse_rows(1:2, valign="top")
```

# Data cleaning
## Median CV coverage vs. mapped duplicate reads

Compare the median coefficient of variation (CV) coverage (`median_cv_coverage`) and percent alignment of reads with duplicates (`mapped_reads_w_dups`).

```{r echo=FALSE, fig.height=4, fig.width=8.5}
# Plot libraries
meta %>% 
  #Create color group for total sequences
  mutate(col.group = ifelse(total_sequences <= 1e5, "<100,000",
                     ifelse(total_sequences <= 3e5, "<300,000",
                     ifelse(total_sequences <= 5e5, "<500,000",
                            ">500,000")))) %>%

ggplot(aes(median_cv_coverage, mapped_reads_w_dups)) +
  geom_point(aes(color=col.group, shape=experiment), size=2) +
  #Label points outside cutoffs
  geom_text_repel(data=filter(meta, median_cv_coverage > 0.6 | 
                              mapped_reads_w_dups < 0.7),
            aes(label=libID), show.legend = FALSE) +
 #Beautify
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(x="Median CV coverage", y="Mapped reads with duplicates",
       color="Total sequences") +
  lims(x=c(0,1), y=c(0,1)) +
  scale_color_manual(values = c("#cbc9e2","#9e9ac8","#756bb1","#54278f"))
```

## Total aligned counts

Plot total counts per library. Here, we compare our standard cutoff of 500,000 to a less strict cutoff of 100,000, because `r nrow(filter(meta, total_sequences <= 5E5))` pDC libraries would be removed at the standard cutoff.

```{r echo=FALSE, fig.height=4, fig.width=8.5}
meta %>% 
  #Create color group for total sequences
  mutate(col.group = ifelse(total_sequences <= 1e5, "<100,000",
                     ifelse(total_sequences <= 3e5, "<300,000",
                     ifelse(total_sequences <= 5e5, "<500,000",
                            ">500,000")))) %>%

ggplot(aes(x=reorder_within(libID, by=total_sequences,
                            within=experiment), 
           y=total_sequences, fill=col.group))  +
  geom_col() +
  # Facet by cell type
  facet_grid(~experiment, scales="free_x", space="free") +
  # Add cutoff line
  geom_hline(yintercept = 1E5) +
  geom_hline(yintercept = 5E5, lty=2) +
  # Beautify
  theme_classic() +
  theme(axis.text.x = element_text(size=rel(0.75),
                                   angle = 90, hjust = 1),
        legend.position = "bottom") +
  labs(x="Library", y="Total aligned counts\n(Log scale)", fill="") +
  scale_x_reordered() +
  scale_fill_manual(values = c("#cbc9e2","#9e9ac8","#756bb1","#54278f")) +
  scale_y_continuous(trans = 'log10',
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)))
```

## Filter by quality and coverage

Thus, the following libraries with questionable metrics (highlighted in red in HTML format) may be removed.

```{r echo=FALSE}
to_remove <- meta %>% 
  filter(total_sequences <= 5E5 | 
         median_cv_coverage > 0.6 |
         mapped_reads_w_dups < 0.75)

to_remove %>% 
  mutate(median_cv_coverage = cell_spec(median_cv_coverage, "html",
                              color = ifelse(median_cv_coverage > 0.6,
                                             "red", "darkgrey")),
         mapped_reads_w_dups = cell_spec(mapped_reads_w_dups, "html",
                              color = ifelse(mapped_reads_w_dups < 0.75,
                                             "red", "darkgrey")),
         total_sequences = cell_spec(total_sequences, "html",
                              color = ifelse(total_sequences <= 1E5,
                                             "red", "darkgrey"))) %>%
  select(experiment, libID, donorID, asthma,IL5,virus,
      median_cv_coverage, mapped_reads_w_dups, total_sequences) %>% 
  arrange(libID) %>%

  kable(align="c", format = "html", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

First, 3 libraries with < 100,000 sequences are removed. Since this removes the media (none:none) sample for donor AC1, all other AC1 samples are also removed. Of note, if the other AC1 samples are retained, they appears as PCA outliers in the next step.

```{r}
#Metadata
meta.filter <- meta %>% 
  # CV coverage and alignment
  filter(total_sequences >= 1E5) %>% 
  filter(donorID != "AC1") %>% 
  droplevels()

#Count data
count.pDC.filter <- count.pDC %>% 
  # Keep libraries (columns) remaining filtered metadata table (rows)
  dplyr::select(geneName, as.character(meta.filter$libID))
```

## Filter PCA outliers

Next, samples are assessed in PCA to determine if remaining questionable quality samples (labeled with libID) appear as PCA outliers.

```{r echo=FALSE, warning=FALSE, message=FALSE}
PCA.all <- count.pDC.filter %>% 
  column_to_rownames("geneName") %>% 
  #Convert to log counts per million
  cpm(., log=TRUE) %>% 
  t() %>% 
  #Calc PCA
  prcomp()

#Extract axes labels with percent contribution of PC
PC1.label <- paste("PC1 (", summary(PCA.all)$importance[2,1]*100, "%)", sep="")
PC2.label <- paste("PC2 (", summary(PCA.all)$importance[2,2]*100, "%)", sep="")

# Extract PC values
PCA.dat <- as.data.frame(PCA.all$x) %>% 
    rownames_to_column("libID") %>%
    # Select PCs for plotting
    dplyr::select(libID, PC1:PC3) %>% 
    # Merge with metadata
    left_join(meta.filter, by="libID")

PCA1 <- ggplot(PCA.dat, aes(PC1, PC2, color=virus:IL5)) +
      geom_point(size=2, stroke=1.5) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nlogCPM") +
      coord_fixed(ratio=1) +
      scale_shape_manual(values=c(21,16)) +
      guides(color=guide_legend(title.position="top", 
                                title.hjust = 0.5))+
      scale_color_manual(values = group.cols)

PCA2 <- ggplot(PCA.dat, aes(PC1, PC2, color=donorID)) +
      geom_point(size=2, stroke=1.5) +
      geom_text_repel(data=filter(PCA.dat, 
                                  libID %in% to_remove$libID),
                      aes(label=libID), show.legend = FALSE) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nlogCPM") +
      coord_fixed(ratio=1) +
      scale_color_manual(values=samp.cols)
```

```{r PCA.outlier, echo=FALSE, message=FALSE}
PCA.all.sd <- PCA.dat %>% 
  group_by(virus, IL5) %>% 
  #Calculate PC mean std deviation
  summarise(.groups="keep",
    PC1.mean = mean(PC1),
    PC1.sd = sd(PC1),
    PC2.mean = mean(PC2),
    PC2.sd = sd(PC2)) %>% 
  #Calculate +/- 3 sd limits
  mutate(
    PC1.min = PC1.mean-(3*PC1.sd),
    PC1.max = PC1.mean+(3*PC1.sd),
    PC2.min = PC2.mean-(3*PC2.sd),
    PC2.max = PC2.mean+(3*PC2.sd))

PCA.all.dat <- PCA.dat %>%   
  full_join(PCA.all.sd) %>% 
  #ID potential outliers
  mutate(col.group = ifelse(PC1 > PC1.max | PC1 < PC1.min |
                            PC2 > PC2.max | PC2 < PC2.min, 
                            "potential outlier", "okay"))

PCA3 <- PCA.all.dat %>%       
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color=col.group),size=2) + 
  geom_text_repel(data=filter(PCA.all.dat,
                            col.group == "potential outlier"),
            aes(label=libID), show.legend = FALSE) +
  #Beautify
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  labs(x=PC1.label, y=PC2.label, 
       title="pDC experiment\nlogCPM") +
  coord_fixed(ratio=1) +
  scale_color_manual(values = c("#969696","#b10026"))
```

```{r echo=FALSE, fig.height=8, fig.width=8.5}
plot_grid(PCA1,PCA2,PCA3, align = "v", ncol=1, axis="l")
```

Donor AC2 appears to be an outlier for all of its samples and will be removed.

```{r}
#Metadata
meta.filter2 <- meta.filter %>% 
  filter(donorID != "AC2") %>% 
  droplevels()

#Count data
count.pDC.filter2 <- count.pDC.filter %>% 
  # Keep libraries (columns) remaining filtered metadata table (rows)
  dplyr::select(geneName, as.character(meta.filter2$libID))
```

Re-assess PCA.

```{r echo=FALSE, warning=FALSE, message=FALSE}
PCA.all <- count.pDC.filter2 %>% 
  column_to_rownames("geneName") %>% 
  #Convert to log counts per million
  cpm(., log=TRUE) %>% 
  t() %>% 
  #Calc PCA
  prcomp()

#Extract axes labels with percent contribution of PC
PC1.label <- paste("PC1 (", summary(PCA.all)$importance[2,1]*100, "%)", sep="")
PC2.label <- paste("PC2 (", summary(PCA.all)$importance[2,2]*100, "%)", sep="")

# Extract PC values
PCA.dat <- as.data.frame(PCA.all$x) %>% 
    rownames_to_column("libID") %>%
    # Select PCs for plotting
    dplyr::select(libID, PC1:PC3) %>% 
    # Merge with metadata
    left_join(meta.filter2, by="libID")

PCA1 <- ggplot(PCA.dat, aes(PC1, PC2, color=virus:IL5)) +
      geom_point(size=2, stroke=1.5) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nlogCPM") +
      coord_fixed(ratio=1) +
      scale_shape_manual(values=c(21,16)) +
      guides(color=guide_legend(title.position="top", 
                                title.hjust = 0.5))+
      scale_color_manual(values = group.cols)

PCA2 <- ggplot(PCA.dat, aes(PC1, PC2, color=donorID)) +
      geom_point(size=2, stroke=1.5) +
      geom_text_repel(data=filter(PCA.dat, 
                                  libID %in% to_remove$libID),
                      aes(label=libID), show.legend = FALSE) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nlogCPM") +
      coord_fixed(ratio=1) +
      scale_color_manual(values=samp.cols)
```

```{r PCA.outlier2, echo=FALSE, message=FALSE}
PCA.all.sd <- PCA.dat %>% 
  group_by(virus, IL5) %>% 
  #Calculate PC mean std deviation
  summarise(.groups="keep",
    PC1.mean = mean(PC1),
    PC1.sd = sd(PC1),
    PC2.mean = mean(PC2),
    PC2.sd = sd(PC2)) %>% 
  #Calculate +/- 3 sd limits
  mutate(
    PC1.min = PC1.mean-(3*PC1.sd),
    PC1.max = PC1.mean+(3*PC1.sd),
    PC2.min = PC2.mean-(3*PC2.sd),
    PC2.max = PC2.mean+(3*PC2.sd))

PCA.all.dat <- PCA.dat %>%   
  full_join(PCA.all.sd) %>% 
  #ID potential outliers
  mutate(col.group = ifelse(PC1 > PC1.max | PC1 < PC1.min |
                            PC2 > PC2.max | PC2 < PC2.min, 
                            "potential outlier", "okay"))

PCA3 <- PCA.all.dat %>%       
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color=col.group),size=2) + 
  geom_text_repel(data=filter(PCA.all.dat,
                            col.group == "potential outlier"),
            aes(label=libID), show.legend = FALSE) +
  #Beautify
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  labs(x=PC1.label, y=PC2.label, 
       title="pDC experiment\nlogCPM") +
  coord_fixed(ratio=1) +
  scale_color_manual(values = c("#969696","#b10026"))
```

```{r echo=FALSE, fig.height=10, fig.width=8.5}
plot_grid(PCA1,PCA2,PCA3, align = "hv", ncol=2, axis="l")
```

Additionally, lib44585 was flagged for poor-quality above and is labeled in the PCA. While this sample is somewhat of an outlier in PCA, it is not > 3 s.d. away from its group's mean. Thus, it will remain and be re-assessed after gene filtering.

Therefore, the filtered data have

* max median CV coverage = `r round(max(meta.filter2$median_cv_coverage), digits=2)`
* min mapped reads with duplicates = `r round(min(meta.filter2$mapped_reads_w_dups), digits=2)`
* min total sequences = `r scales::comma_format()(min(meta.filter2$total_sequences))`

## Normalize for RNA composition

Create DGEList object

```{r}
dat.pDC.filter2 <- DGEList(
  #count table. move gene names to column names
  counts=as.matrix(column_to_rownames(count.pDC.filter2, "geneName")),
  #metadata
  samples=meta.filter2,
  #keep genes in count table
  genes=filter(key.pc, geneName %in% count.pDC.filter2$geneName))
```

## Filter rare genes

The raw gene sets contain highly variable, low abundance/rare genes. Genes must be at least 1 CPM in at least 3 samples to remain.

```{r}
source("https://raw.githubusercontent.com/kdillmcfarland/R_bioinformatic_scripts/master/RNAseq_rare_gene_filter.R")

rare.gene.filter(dat.pDC.filter2, 
                 min.CPM=1,
                 min.sample=3, 
                 name="dat.pDC.filter2.abund")
```

```{r echo=FALSE, warning=FALSE}
#Calc normalization factors
dat.pDC.filter2.norm <- calcNormFactors(dat.pDC.filter2)

#voom normalize
temp <- voomWithQualityWeights(dat.pDC.filter2.norm, 
     design=model.matrix(~ IL5 + virus,
                         data=dat.pDC.filter2.norm$samples),
     plot=FALSE, save.plot = TRUE)

MV1 <- data.frame(
  x = temp$voom.xy$x, 
  y = temp$voom.xy$y,
  linex = temp$voom.line$x, 
  liney = temp$voom.line$y) %>%

  ggplot() +
  geom_point(aes(x=x, y=y), size=0.5) +
  geom_path(aes(x=linex, y=liney), color="red") +
  theme_classic() +
  labs(x="log2( count size + 0.5 )", y="Sqrt (stdev)",
       title="All genes")
```

```{r echo=FALSE}
#Calc normalization factors
dat.pDC.filter2.abund.norm <- calcNormFactors(dat.pDC.filter2.abund)

#Voom normalize
dat.pDC.filter2.abund.norm.voom <- voomWithQualityWeights(dat.pDC.filter2.abund.norm, 
     design=model.matrix(~ IL5 + virus,
            data=dat.pDC.filter2.abund.norm$samples),
     plot=FALSE, save.plot = TRUE)

#plot mean-var trend
MV2 <- data.frame(
  x = dat.pDC.filter2.abund.norm.voom$voom.xy$x, 
  y = dat.pDC.filter2.abund.norm.voom$voom.xy$y,
  linex = dat.pDC.filter2.abund.norm.voom$voom.line$x, 
  liney = dat.pDC.filter2.abund.norm.voom$voom.line$y) %>%

  ggplot() +
  geom_point(aes(x=x, y=y), size=0.5) +
  geom_path(aes(x=linex, y=liney), color="red") +
  theme_classic() +
  labs(x="log2( count size + 0.5 )", y="Sqrt (stdev)",
       title="Abundant genes")
```

```{r echo=FALSE, fig.height=4, fig.width=8.5}
plot_grid(MV1, MV2)
```

```{r echo=FALSE}
genes <- nrow(dat.pDC.filter2.norm$genes)
genes.abund <- nrow(dat.pDC.filter2.abund.norm$genes)
```

This removes `r genes-genes.abund` (~ `r round((genes-genes.abund)/genes*100, digits=0)`%) genes. This sufficiently filters these data as indicated by a decreasing variance trend for lowly abundant genes.

## PCA

We see that previously questionable samples (labeled with libID) are sufficiently normalized to not appear as PCA outliers.

```{r echo=FALSE, warning=FALSE, message=FALSE}
PCA.all <- as.data.frame(dat.pDC.filter2.abund.norm.voom$E) %>% 
  t() %>% 
  #Calc PCA
  prcomp()

#Extract axes labels with percent contribution of PC
PC1.label <- paste("PC1 (", summary(PCA.all)$importance[2,1]*100, "%)", sep="")
PC2.label <- paste("PC2 (", summary(PCA.all)$importance[2,2]*100, "%)", sep="")

# Extract PC values
PCA.dat <- as.data.frame(PCA.all$x) %>% 
    rownames_to_column("libID") %>%
    # Select PCs for plotting
    dplyr::select(libID, PC1:PC3) %>% 
    # Merge with metadata
    left_join(meta.filter2, by="libID")

PCA1 <- ggplot(PCA.dat, aes(PC1, PC2, color=virus:IL5)) +
      geom_point(size=2, stroke=1.5) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nvoom normalized abundant logCPM") +
      coord_fixed(ratio=1) +
      scale_shape_manual(values=c(21,16)) +
      guides(color=guide_legend(title.position="top", 
                                title.hjust = 0.5))+
      scale_color_manual(values = group.cols)

PCA2 <- ggplot(PCA.dat, aes(PC1, PC2, color=donorID)) +
      geom_point(size=2, stroke=1.5) +
      geom_text_repel(data=filter(PCA.dat, 
                                  libID %in% to_remove$libID),
                      aes(label=libID), show.legend = FALSE) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nvoom normalized abundant logCPM") +
      coord_fixed(ratio=1) +
      scale_color_manual(values=samp.cols)
```

```{r PCA.outlier3, echo=FALSE, message=FALSE}
PCA.all.sd <- PCA.dat %>% 
  group_by(virus, IL5) %>% 
  #Calculate PC mean std deviation
  summarise(.groups="keep",
    PC1.mean = mean(PC1),
    PC1.sd = sd(PC1),
    PC2.mean = mean(PC2),
    PC2.sd = sd(PC2)) %>% 
  #Calculate +/- 3 sd limits
  mutate(
    PC1.min = PC1.mean-(3*PC1.sd),
    PC1.max = PC1.mean+(3*PC1.sd),
    PC2.min = PC2.mean-(3*PC2.sd),
    PC2.max = PC2.mean+(3*PC2.sd))

PCA.all.dat <- PCA.dat %>%   
  full_join(PCA.all.sd) %>% 
  #ID potential outliers
  mutate(col.group = ifelse(PC1 > PC1.max | PC1 < PC1.min |
                            PC2 > PC2.max | PC2 < PC2.min, 
                            "potential outlier", "okay"))

PCA3 <- PCA.all.dat %>%       
  ggplot(aes(PC1, PC2)) +
  geom_point(aes(color=col.group),size=2) + 
  geom_text_repel(data=filter(PCA.all.dat,
                            col.group == "potential outlier"),
            aes(label=libID), show.legend = FALSE) +
  #Beautify
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  labs(x=PC1.label, y=PC2.label, 
       title="pDC experiment\nvoom normalized abundant logCPM") +
  coord_fixed(ratio=1) +
  scale_color_manual(values = c("#969696","#b10026"))
```

```{r echo=FALSE, fig.height=10, fig.width=8.5}
plot_grid(PCA1,PCA2,PCA3, align = "hv", ncol=2, axis="l")
```

## Summary

The following final quality cutoffs were applied to the data:

* Library median CV coverage $\leq$ `r round(max(dat.pDC.filter2.abund.norm.voom$targets$median_cv_coverage), digits=2)`
* Library mapped reads with duplicates $\geq$ `r round(min(dat.pDC.filter2.abund.norm.voom$targets$mapped_reads_w_dups), digits=2)`
* Library total counts $\geq$ `r min(dat.pDC.filter2.abund.norm.voom$targets$total_sequences)`
* Libraries from donor AC1 due to loss of media sample in previous filtering
* Within 3 sd of group mean on PCA (*e.g.* not outliers)
* Genes at least 1 CPM in at least 3 samples

```{r echo=FALSE, fig.height=4, fig.width=8.5}
# Plot libraries
meta %>% 
  #Create color group for removed samples
  mutate(col.group = ifelse(libID %in% dat.pDC.filter2$samples$libID,
                            "Retained","Removed")) %>%

ggplot(aes(median_cv_coverage, mapped_reads_w_dups)) +
  geom_point(aes(color=col.group, shape=experiment), size=2) +
  geom_hline(yintercept = 0.85, lty=2) +
  geom_vline(xintercept = 0.85, lty=2) +
  scale_y_continuous(breaks = sort(c(seq(0,1,length.out=5), 0.85)),
                    limits = c(0,1)) +
  scale_x_continuous(breaks = sort(c(seq(0,1, length.out=5), 0.85)),
                     limits = c(0,1)) +
  #Beautify
  theme_classic() +
  theme(legend.position = "bottom") +
  labs(x="Median CV coverage", y="Mapped reads with duplicates",
       color="Total sequences",
       title="Median CV coverage vs. mapped duplicate reads")
```

```{r echo=FALSE, fig.height=4, fig.width=8.5}
meta %>% 
  #Create color group for removed samples
  mutate(col.group = ifelse(libID %in% dat.pDC.filter2$samples$libID,
                            "Retained","Removed")) %>%

ggplot(aes(x=reorder_within(libID, by=total_sequences,
                            within=experiment), 
           y=total_sequences, fill=col.group))  +
  geom_col() +
  # Facet by cell type
  facet_grid(~experiment, scales="free_x", space="free") +
  # Add cutoff line
  geom_hline(yintercept = 1E5, lty=2) +
  # Beautify
  theme_classic() +
  theme(axis.text.x = element_text(size=rel(0.75),
                                   angle = 90, hjust = 1),
        legend.position = "bottom") +
  labs(x="Library", y="Total aligned counts\n(Log scale)", fill="",
       title="Total aligned counts") +
  scale_x_reordered() +
  scale_y_continuous(trans = 'log10',
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)))
```

```{r echo=FALSE}
filter.dup <- meta %>% 
  filter(mapped_reads_w_dups <= 0.85) %>% 
  mutate(filter = "mapped dup reads")
filter.cv <- meta %>% 
  filter(median_cv_coverage >= 0.85)%>% 
  mutate(filter = "CV coverage")
filter.seq <- meta %>% 
  filter(total_sequences <= 1E5)%>% 
  mutate(filter = "total seqs")
filter.AC1 <- meta %>% 
  filter(donorID == "AC1" & virus != "none")%>% 
  mutate(filter = "no media-only sample")
filter.AC2 <- meta %>% 
  filter(donorID == "AC2")%>% 
  mutate(filter = "PCA outlier")

bind_rows(filter.dup,filter.cv,filter.seq,
          filter.AC1,filter.AC2) %>% 
  group_by(libID) %>% 
  mutate(filters= paste(filter, collapse = ", \n")) %>% 
  select(experiment, libID, donorID, 
         asthma,IL5,virus, filters) %>%
  distinct() %>% 
  arrange(experiment, donorID, asthma,IL5,virus) %>%

  kable(align=c("c","c","c","c","l")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)  
```

This results in the following samples for statistical analysis.

```{r echo=FALSE}
meta.filter2 %>% 
  group_by(experiment, asthma,IL5,virus) %>% 
  tally() %>% 
  arrange(experiment, asthma,IL5,virus) %>%

  kable(align="c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>% 
  collapse_rows(1:2, valign="top") 
```

# Save clean data

Normalized dat objects.

```{r message=FALSE}
#Save R object
#Rename all pDC object
dat.pDC.voom <- dat.pDC.filter2.abund.norm.voom

#P259_1
dat.pDC.voom_1 <- dat.pDC.voom

dat.pDC.voom_1$targets <- dat.pDC.voom_1$targets %>% 
  rownames_to_column() %>% 
  filter(experiment == "P259_1") %>% 
  droplevels() %>% 
  column_to_rownames()

dat.pDC.voom_1$E <- as.data.frame(dat.pDC.voom_1$E) %>% 
  rownames_to_column() %>% 
  select(rowname, dat.pDC.voom_1$targets$libID) %>% 
  column_to_rownames()

#P259_2
dat.pDC.voom_2 <- dat.pDC.voom

dat.pDC.voom_2$targets <- dat.pDC.voom_2$targets %>% 
  rownames_to_column() %>% 
  filter(experiment == "P259_2") %>% 
  droplevels() %>% 
  column_to_rownames()

dat.pDC.voom_2$E <- as.data.frame(dat.pDC.voom_2$E) %>% 
  rownames_to_column() %>% 
  select(rowname, dat.pDC.voom_2$targets$libID) %>% 
  column_to_rownames()

save(dat.pDC.voom, dat.pDC.voom_1, dat.pDC.voom_2,
     file="data_clean/P259_pDC_clean.RData")
```

# R session

```{r}
sessionInfo()
```

***