### 2020-09-09 2.P259.pDC_model.selection.Rmd

---
title: 'P259: Model selection'
subtitle: "Dendritic cells (pDC)"
author: "Kim Dill-McFarland, kadm@uw.edu"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: console
---
# Background

The purpose of this workflow is to identify differentially expressed genes in pDC. Several models are compared in order to determine the best way to extract the effects of Anti-IL5 therapy and eosinophil (EOS) supernatant over the much larger human rhinovirus (RV) signal.

# Setup
Load packages

```{r message=FALSE, warning=FALSE}
# Data manipulation and figures
library(tidyverse)
    # Multi-panel figures for ggplot
    library(cowplot)
    # Add p-values to plots
    library(ggpubr)
#Venn diagrams
library(venn)
#Linear models
library(limma)
#Define ggplot colors
group.cols <- c("none:none"="#dadaeb",
               "none:AntiIL5"="#9e9ac8",
               "none:EOS.supp"="#54278f",
               "HRV:none"="#c7e9c0",
               "HRV:AntiIL5"="#74c476",
               "HRV:EOS.supp"="#006d2c",
               "flu:none"="#fdae6b",
               "flu:AntiIL5"="#e6550d")
samp.cols <- c("AC1"="#969696",
               "AC2"="#a6cee3",
               "AC3"="#1f78b4",
               "AC4"="#b2df8a",
               "AC5"="#33a02c",
               "AT1"="#fb9a99",
               "AT2"="#e31a1c",               
               "AT3"="#fdbf6f",
               "AT4"="#ff7f00",
               "donor1"="#cab2d6",
               "donor2"="#6a3d9a",
               "donor3"="#ffff99",
               "donor4"="#b15928")

# Print tty table to knit file
library(knitr)
library(kableExtra)
```

Set seed

```{r}
set.seed(4389)
```

Scripts and functions

```{r}
#To extract and format p-value from lmFit
source("https://raw.githubusercontent.com/kdillmcfarland/R_bioinformatic_scripts/master/limma.extract.pval.R")
#opposite of %in%
`%notin%` <- Negate(`%in%`)
```

# Load data

Load normalized edgeR data objects. RNA-seq data was cleaned and normalized in [data cleaning](1.P259.pDC_data.cleaning.html). This included filtering for median CV coverage, mapped duplicate reads, total aligned counts, PCA outliers, and rare genes.

```{r}
load("data_clean/P259_pDC_clean.RData")
```

This includes in the following pDC samples.

```{r echo=FALSE}
dat.pDC.voom$targets %>% 
  group_by(experiment, asthma,IL5,virus) %>% 
  tally() %>% 
  arrange(experiment, asthma,IL5,virus) %>%

  kable(align="c") %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>% 
  collapse_rows(1:2, valign="top") 
```

# Variables of interest

**Main effects**

* `donorID`: `r class(dat.pDC.voom$targets$donorID)` for each donor (N = `r length(unique(dat.pDC.voom$targets$donorID))`)
* `IL5`: `r class(dat.pDC.voom$targets$IL5)` of patient AntiIL5 therapy or sample EOS supernatant treatment (`r paste(levels(dat.pDC.voom$targets$IL5), collapse=", ")`)
* `virus`: `r class(dat.pDC.voom$targets$virus)` of viral conditions (`r paste(levels(dat.pDC.voom$targets$virus), collapse=", ")`)

**Fixed effects**

* `Sex`: `r paste(unique(dat.pDC.voom$targets$Sex), collapse=", ")`
* `Age`: age, years (`r min(dat.pDC.voom$targets$Age, na.rm=TRUE)` - `r max(dat.pDC.voom$targets$Age, na.rm=TRUE)`)
    - Neither available for 2 of 4 donors in P259.1

## Explore fixed effects and possible confounding variables

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8.5}
plot1 <- dat.pDC.voom$targets %>% 
  select(experiment,donorID,IL5,Age) %>% 
  distinct() %>%

  ggplot(aes(x=IL5, y=Age)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color=donorID), height=0,width=0.2, size=3) +
  stat_compare_means(method="wilcox.test") +
  facet_wrap(~experiment, scales = "free_x") +
  theme_classic() +
  scale_color_manual(values=samp.cols) +
  labs(x="Treatment")

#Calculate chi-square for sex
sex.p.1 <- dat.pDC.voom$targets %>% 
  filter(experiment == "P259_1") %>% 
  droplevels()
sex.p.1 <- table(sex.p.1$IL5, sex.p.1$Sex) %>% 
  chisq.test()

sex.p.2 <- dat.pDC.voom$targets %>% 
  filter(experiment == "P259_2") %>% 
  droplevels()
sex.p.2 <- table(sex.p.2$IL5, sex.p.2$Sex) %>% 
  chisq.test()

plot2 <- dat.pDC.voom$targets %>% 
  group_by(experiment, IL5, Sex) %>% 
  tally() %>%

  ggplot(aes(x=IL5, y=n, fill=Sex)) +
  geom_bar(position="fill", stat="identity") +
  facet_wrap(~experiment, scales = "free_x") +
  theme_classic() +
  labs(y="Proportion", title=paste("Sex\nP259.1 Chi-squared = ", 
                                   round(sex.p.1$p.value,3),
                                   "\nP259.2 Chi-squared = ", 
                                   round(sex.p.2$p.value,3),
                                   sep="")) +
  scale_fill_manual(values=c("#f1a340","#998ec3"), na.value="grey")+
  labs(x="Treatment")

plot_grid(plot1, plot2) 
```

Age and sex appear to be different between treatment groups but are technically not significantly different, likely as the result of the small sample size. Due to this, the small sample size, and missing data in P259.1, sex and age will not be used as co-variates.

# PCA

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=5, fig.width=8.5}
# Calculate PCA
PCA <- as.data.frame(dat.pDC.voom$E) %>% 
  t() %>% 
  prcomp()

PC1.label <- paste("PC1 (", summary(PCA)$importance[2,1]*100, "%)", sep="")
PC2.label <-paste("PC2 (", summary(PCA)$importance[2,2]*100, "%)", sep="")

# Extract PC values
PCA.dat <- as.data.frame(PCA$x) %>% 
  rownames_to_column("libID") %>%
  # Select PCs
  dplyr::select(libID, PC1:PC3) %>% 
  # Merge with metadata
  left_join(dat.pDC.voom$targets, by="libID")

PCA1 <- ggplot(PCA.dat, aes(PC1, PC2, color=virus:IL5, shape=experiment)) +
      geom_point(size=2, stroke=1.5)  +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nvoom normalized abundant logCPM") +
      coord_fixed(ratio=1) +
      scale_shape_manual(values=c(21,16)) +
      guides(color=guide_legend(title.position="top", 
                                title.hjust = 0.5))+
      scale_color_manual(values = group.cols)

PCA2 <- ggplot(PCA.dat, aes(PC1, PC2, color=donorID, shape=experiment)) +
      geom_point(size=2, stroke=1.5) +
      #Beautify
      theme_classic() +
      labs(x=PC1.label, y=PC2.label, 
           title="pDC experiment\nvoom normalized abundant logCPM") +
      coord_fixed(ratio=1) +
      scale_shape_manual(values=c(21,16)) +
      scale_color_manual(values=samp.cols)

plot_grid(PCA1,PCA2, align = "hv", ncol=2)
```

# P259.1 model comparison
## Run models
### Interaction model

`gene expression ~ EOS.supp * virus + (1|donor)`

```{r}
# Define model
model_1.interact <- model.matrix(~ IL5*virus,
                          data=dat.pDC.voom_1$targets)
    colnames(model_1.interact) <- c("(Intercept)", "EOS.supp",
                            "HRV", "EOS.supp:HRV")
#Block by donor
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_1$E,
  model_1.interact,
  block=dat.pDC.voom_1$targets$donorID)$consensus.correlation

consensus.corr

# Fit model to transformed count data. Calculate eBayes
efitQW_1.interact <- eBayes(
                lmFit(dat.pDC.voom_1$E, 
                      model_1.interact,
                      block=dat.pDC.voom_1$targets$donorID,
                      correlation=consensus.corr))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_1.interact,
             voom.dat=dat.pDC.voom_1$E, 
             eFit=efitQW_1.interact, 
             name="pval_1.interact",
             summary=TRUE,
             contrasts=FALSE)
```

Summary

```{r echo=FALSE}
pval_1.interact.summ  %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

### Contrasts model

`gene expression ~ EOS.supp:virus + (1|donor)`

```{r}
# Define model
dat.pDC.voom_1$targets <- dat.pDC.voom_1$targets %>% 
  mutate(contrast=paste(IL5, virus, sep="_")) %>% 
  mutate(contrast=factor(contrast, 
                         levels = c("none_none","none_HRV",
                                 "EOS.supp_none","EOS.supp_HRV")))

model_1.contrast<- model.matrix(~ 0 + contrast,
                          data=dat.pDC.voom_1$targets)
    colnames(model_1.contrast) <- c(
                                 "none_none","none_HRV",
                                 "EOS.supp_none","EOS.supp_HRV")
#Block by donor  
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_1$E,
  model_1.contrast,
  block=dat.pDC.voom_1$targets$donorID)$consensus.correlation

consensus.corr

#Fit model
fitQW_1.contrast <- lmFit(dat.pDC.voom_1$E, 
                      model_1.contrast,
                      block=dat.pDC.voom_1$targets$donorID,
                      correlation=consensus.corr)
#Get contrasts
contrast.matrix <- makeContrasts(
  none_HRV-none_none, 
  EOS.supp_HRV-EOS.supp_none,
  EOS.supp_none-none_none,
  EOS.supp_HRV-none_HRV,
  levels=model_1.contrast)

efitQW_1.contrast <- eBayes(contrasts.fit(fitQW_1.contrast,
                                  contrast.matrix))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_1.contrast,
             voom.dat=dat.pDC.voom_1$E, 
             eFit=efitQW_1.contrast, 
             name="pval_1.contrast",
             summary=TRUE,
             contrasts=TRUE,
             contrast.mat=contrast.matrix)
```

Summary

```{r echo=FALSE}
pval_1.contrast.summ  %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

#### Contrast selected genes

Genes of interest are those that change with virus (*e.g.* significant for virus in untreated and/or treated samples) AND with EOS supernatant in media OR HRV.

```{r echo=FALSE, fig.height=6, fig.width=8.5}
#Define FDR cutoffs to assess
fdr.list <- c(0.05,0.1,0.2,0.3,0.4,0.5)

par(mfrow = c(2, 3))

for(fdr in fdr.list){
  #Blank list for results
  venn.list <- list()
  #Signif for virus
  temp <- pval_1.contrast %>% 
          filter(group %in% c("none_HRV - none_none",
                              "EOS.supp_HRV - EOS.supp_none") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["RV-infected\nvs media"]] <- temp

  #Signif in media
  temp2 <- pval_1.contrast %>% 
          filter(group %in% c("EOS.supp_none - none_none") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)

  venn.list[["- RV\nEOS supernatant\nvs none"]] <- temp2

  #Signif in HRV
  temp3 <- pval_1.contrast %>% 
          filter(group %in% c("EOS.supp_HRV - none_HRV") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)

  venn.list[["+ RV\nEOS supernatant\nvs none"]] <- temp3

  venn(ilab=FALSE, zcolor = "style",ilcs=1.5,sncs=1.5,
     x=venn.list, box=FALSE)

  title(sub=paste("FDR < ", fdr, sep=""), line = -1, cex.sub=1.5)
}
```

### Delta model

`RV - media gene expression ~ EOS.supp + (1|donor)`

Calculate HRV - media counts.

```{r echo=FALSE, message=FALSE}
dat.pDC.voom_1.delta.E <- as.data.frame(dat.pDC.voom_1$E) %>% 
  #Longer format, libID to columns
  rownames_to_column() %>% 
  pivot_longer(-rowname, names_to="libID") %>% 
  #Merge with metadata
  left_join(select(dat.pDC.voom_1$targets, libID, 
                   donorID, IL5, virus), by = "libID") %>% 
  select(-libID) %>% 
  #Separate media vs HRV samples
  pivot_wider(names_from = virus) %>% 
  #Calculate delta
  group_by(rowname, donorID, IL5) %>% 
  summarise(value = HRV-none) %>% 
  drop_na(value) %>% 
  ungroup() %>% 
  #Rename with donor.EOS
  mutate(new.name = paste(donorID, IL5, sep="_")) %>% 
  select(-donorID, -IL5) %>% 
  arrange(new.name, rowname) %>% 
  pivot_wider(names_from = new.name) %>% 
  #Matrix
  column_to_rownames()
```

```{r echo=FALSE}
#Subset metadata
dat.pDC.voom_1.delta.tgts <- dat.pDC.voom_1$targets %>% 
  select(-lib.size, -norm.factors, -libID, -sample.weights) %>% 
  filter(virus=="HRV") %>% 
  mutate(new.name = paste(donorID, IL5, sep="_")) %>% 
  filter(new.name %in% colnames(dat.pDC.voom_1.delta.E)) %>% 
  arrange(new.name)
```

Fit model

```{r}
# Define model
model_1.delta <- model.matrix(~ IL5,
                          data=dat.pDC.voom_1.delta.tgts)
    colnames(model_1.delta) <- c("(Intercept)","EOS.supp")

#Block by donor
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_1.delta.E,
  model_1.delta,
  block=dat.pDC.voom_1.delta.tgts$donorID)$consensus.correlation

consensus.corr

# Fit model to transformed count data. Calculate eBayes
efitQW_1.delta <- eBayes(
                lmFit(dat.pDC.voom_1.delta.E, 
                      model_1.delta,
                      block=dat.pDC.voom_1.delta.tgts$donorID,
                      correlation=consensus.corr))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_1.delta,
             voom.dat=dat.pDC.voom_1.delta.E, 
             eFit=efitQW_1.delta, 
             name="pval_1.delta",
             summary=TRUE,
             contrasts=FALSE)
```

Summary

```{r echo=FALSE}
pval_1.delta.summ %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

## Compare models

```{r echo=FALSE}
fdr.cutoff <- 0.2
```

Total genes significant in one or more models at FDR < `r fdr.cutoff`

```{r echo=FALSE}
venn.list <- list()

#Signif for interaction
  temp <- pval_1.interact %>% 
          filter(group %in% c("EOS.supp:HRV") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif interaction"]] <- temp

#Signif for delta
  temp <- pval_1.delta %>% 
          filter(group %in% c("EOS.supp") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif delta"]] <- temp

#Signif for contrasts
  ##Signif for virus
  temp1 <- pval_1.contrast %>% 
          filter(group %in% c("none_HRV - none_none", 
                              "EOS.supp_HRV - EOS.supp_none") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  ##Signif for treatment
  temp2 <- pval_1.contrast %>% 
          filter(group %in% c("EOS.supp_none - none_none", 
                              "EOS.supp_HRV - none_HRV") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  temp <- pval_1.contrast %>% 
          filter(geneName %in% temp1 & geneName %in% temp2) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif contrast"]] <- temp

par(mfrow = c(1,1)) 
  venn(ilab=FALSE, zcolor = "style",ilcs=1,sncs=1.5,
     x=venn.list)

  title(sub="FDR < 0.2", line = -1, cex.sub=1.5)
```

## Save results

```{r}
dir.create("results/gene_level/P259.1_model.selection/", 
           showWarnings = FALSE, recursive = TRUE)

save(pval_1.interact, pval_1.contrast, pval_1.delta,
     file="results/gene_level/P259.1_model.selection/P259.1_gene_pval.RData")

#gene lists
save(venn.list, file="results/gene_level/P259.1_model.selection/P259.1_gene_lists.RData")

#pval csv
pval_1.interact %>% 
  mutate(model = "interaction") %>% 
  bind_rows(mutate(pval_1.contrast, 
                   model="contrasts")) %>% 
  bind_rows(mutate(pval_1.delta, 
                   model="delta")) %>% 
  #Add HGNC symbol
  left_join(dat.pDC.voom_1$genes) %>% 
  write_csv(path = "results/gene_level/P259.1_gene_pval.csv")
```

# P259.2 model comparison
## Run models
### Interaction model

`gene expression ~ AntiIL5 * virus + (1|donor)`

```{r}
# Define model
model_2.interact <- model.matrix(~ IL5*virus,
                          data=dat.pDC.voom_2$targets)
    colnames(model_2.interact) <- c("(Intercept)", "AntiIL5",
                            "HRV", "AntiIL5:HRV")
#Block by donor
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_2$E,
  model_2.interact,
  block=dat.pDC.voom_2$targets$donorID)$consensus.correlation

consensus.corr

# Fit model to transformed count data. Calculate eBayes
efitQW_2.interact <- eBayes(
                lmFit(dat.pDC.voom_2$E, 
                      model_2.interact,
                      block=dat.pDC.voom_2$targets$donorID,
                      correlation=consensus.corr))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_2.interact,
             voom.dat=dat.pDC.voom_2$E, 
             eFit=efitQW_2.interact, 
             name="pval_2.interact",
             summary=TRUE,
             contrasts=FALSE)
```

Summary

```{r echo=FALSE}
pval_2.interact.summ  %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

### Contrasts model

`gene expression ~ AntiIL5:virus + (1|donor)`

```{r}
# Define model
dat.pDC.voom_2$targets <- dat.pDC.voom_2$targets %>% 
  mutate(contrast=paste(IL5,virus, sep="_")) %>% 
  mutate(contrast=factor(contrast, 
                         levels = c("none_none","none_HRV",
                                 "AntiIL5_none","AntiIL5_HRV")))

model_2.contrast<- model.matrix(~ 0 + contrast,
                          data=dat.pDC.voom_2$targets)
    colnames(model_2.contrast) <- c(
                                 "none_none","none_HRV",
                                 "AntiIL5_none","AntiIL5_HRV")
#Block by donor  
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_2$E,
  model_2.contrast,
  block=dat.pDC.voom_2$targets$donorID)$consensus.correlation

consensus.corr

#Fit model
fitQW_2.contrast <- lmFit(dat.pDC.voom_2$E, 
                      model_2.contrast,
                      block=dat.pDC.voom_2$targets$donorID,
                      correlation=consensus.corr)

#Get contrasts
contrast.matrix <- makeContrasts(
  none_HRV-none_none, 
  AntiIL5_HRV-AntiIL5_none,
  AntiIL5_none-none_none,
  AntiIL5_HRV-none_HRV,
  levels=model_2.contrast)

efitQW_2.contrast <- eBayes(contrasts.fit(fitQW_2.contrast,
                                  contrast.matrix))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_2.contrast,
             voom.dat=dat.pDC.voom_2$E, 
             eFit=efitQW_2.contrast, 
             name="pval_2.contrast",
             summary=TRUE,
             contrasts=TRUE,
             contrast.mat=contrast.matrix)
```

Summary

```{r echo=FALSE}
pval_2.contrast.summ  %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

### Contrast selected genes

Genes of interest are those that change with virus (*e.g.* significant for virus in untreated and/or treated donors) AND are different between untreated and treated donors in media OR RV.

```{r echo=FALSE, fig.height=6, fig.width=8.5}
#Define FDR cutoffs to assess
fdr.list<- c(0.05,0.1,0.2,0.3,0.4,0.5)

par(mfrow = c(2, 3))

for(fdr in fdr.list){
  #Blank list for results
  venn.list <- list()
  #Signif for virus
  temp <- pval_2.contrast %>% 
          filter(group %in% c("none_HRV - none_none",
                              "AntiIL5_HRV - AntiIL5_none") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["RV-infected\nvs media"]] <- temp

  #Signif in media
  temp2 <- pval_2.contrast %>% 
          filter(group %in% c("AntiIL5_none - none_none") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)

  venn.list[["- RV\nAntiIL5 therapy\nvs none"]] <- temp2

  #Signif in HRV
  temp3 <- pval_2.contrast %>% 
          filter(group %in% c("AntiIL5_HRV - none_HRV") & 
                   adj.P.Val <= fdr) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)

  venn.list[["+ RV\nAntiIL5 therapy\nvs none"]] <- temp3

  venn(ilab=FALSE, zcolor = "style",ilcs=1.5,sncs=1.5,
     x=venn.list, box=FALSE)

  title(sub=paste("FDR < ", fdr, sep=""), line = -1, cex.sub=1.5)
}
```

### Delta model

`RV - media gene expression ~ AntiIL5`

Calculate HRV - media counts.

```{r echo=FALSE, message=FALSE}
dat.pDC.voom_2.delta.E <- as.data.frame(dat.pDC.voom_2$E) %>% 
  #Longer format, libID to columns
  rownames_to_column() %>% 
  pivot_longer(-rowname, names_to="libID") %>% 
  #Merge with metadata
  left_join(select(dat.pDC.voom_2$targets, libID, 
                   donorID, virus.detail), by = "libID") %>% 
  select(-libID) %>% 
  #Separate media vs HRV samples
  pivot_wider(names_from = virus.detail) %>% 
  #Calculate delta
  group_by(rowname, donorID) %>% 
  summarise(oldHRV = oldHRV-none,
            newHRV = newHRV-none) %>% 
  #Rename with donor.HRV
  pivot_longer(-c(rowname, donorID)) %>% 
  mutate(new.name = paste(donorID, name, sep=".")) %>% 
  #Wide format
  select(rowname, new.name, value) %>% 
  arrange(new.name, rowname) %>% 
  pivot_wider(names_from = new.name) %>% 
  #Matrix
  column_to_rownames()
```

```{r echo=FALSE}
#Subset metadata
dat.pDC.voom_2.delta.tgts <- dat.pDC.voom_2$targets %>% 
  select(-lib.size, -norm.factors, -libID, -sample.weights) %>% 
  mutate(new.name = paste(donorID, virus.detail, sep=".")) %>% 
  filter(new.name %in% colnames(dat.pDC.voom_2.delta.E)) %>% 
  arrange(new.name)
```

Fit model

```{r}
# Define model
model_2.delta <- model.matrix(~ IL5,
                          data=dat.pDC.voom_2.delta.tgts)
    colnames(model_2.delta) <- c("(Intercept)","AntiIL5")

#Block by donor
consensus.corr <- duplicateCorrelation(
  dat.pDC.voom_2.delta.E,
  model_2.delta,
  block=dat.pDC.voom_2.delta.tgts$donorID)$consensus.correlation

consensus.corr

# Fit model to transformed count data. Calculate eBayes
efitQW_2.delta <- eBayes(
                lmFit(dat.pDC.voom_2.delta.E, 
                      model_2.delta,
                      block=dat.pDC.voom_2.delta.tgts$donorID,
                      correlation=consensus.corr))
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
#Extract p-values from results
extract.pval(model=model_2.delta,
             voom.dat=dat.pDC.voom_2.delta.E, 
             eFit=efitQW_2.delta, 
             name="pval_2.delta",
             summary=TRUE,
             contrasts=FALSE)
```

Summary

```{r echo=FALSE}
pval_2.delta.summ %>%

  kable(align=c("l","c","c","c","c","c","c"),
        col.names = c("Variable", 
                      "<0.05", "<0.1",
                      "<0.2", "<0.3", "<0.4",
                      "<0.5")) %>% 
  kable_styling(latex_options="HOLD_position") %>% 
  add_header_above(c(" ", "Total genes at FDR" = 6))
```

## Compare models

Total genes significant in one or more models at FDR < `r fdr.cutoff`

```{r echo=FALSE}
venn.list <- list()

#Signif for interaction
  temp <- pval_2.interact %>% 
          filter(group %in% c("AntiIL5:HRV") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif interaction"]] <- temp

#Signif for delta
  temp <- pval_2.delta %>% 
          filter(group %in% c("AntiIL5") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif delta"]] <- temp

#Signif for contrasts
  ##Signif for virus
  temp1 <- pval_2.contrast %>% 
          filter(group %in% c("none_HRV - none_none", 
                              "AntiIL5_HRV - AntiIL5_none") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  ##Signif for treatment
  temp2 <- pval_2.contrast %>% 
          filter(group %in% c("AntiIL5_none - none_none", 
                              "AntiIL5_HRV - none_HRV") & 
                   adj.P.Val <= fdr.cutoff) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  temp <- pval_2.contrast %>% 
          filter(geneName %in% temp1 & geneName %in% temp2) %>% 
          distinct(geneName) %>% 
          select(geneName) %>% unlist(use.names = FALSE)
  venn.list[["Signif contrast"]] <- temp

par(mfrow = c(1,1)) 
  venn(ilab=FALSE, zcolor = "style",ilcs=1,sncs=1.5,
     x=venn.list)

  title(sub="FDR < 0.5", line = -1, cex.sub=1.5)
```

## Save results

```{r}
dir.create("results/gene_level/P259.2_model.selection/", 
           showWarnings = FALSE, recursive = TRUE)

save(pval_2.interact, pval_2.contrast, pval_2.delta,
     file="results/gene_level/P259.2_model.selection/P259.2_gene_pval.RData")

#gene lists
save(venn.list, file="results/gene_level/P259.2_model.selection/P259.2_gene_lists.RData")

#pval csv
pval_2.interact %>% 
  mutate(model = "interaction") %>% 
  bind_rows(mutate(pval_2.contrast, 
                   model="contrasts")) %>% 
  bind_rows(mutate(pval_2.delta, 
                   model="delta")) %>% 
  #Add HGNC symbol
  left_join(dat.pDC.voom_2$genes) %>% 
  write_csv(path = "results/gene_level/P259.2_gene_pval.csv")
```

# R session

```{r}
sessionInfo()
```

***