### 2021-05-10 publication/4.P259.pDC_publication.fig.table.Rmd
Mv pub rmd to publication dir

---
title: 'P259: Publication figures and tables'
subtitle: "Dendritic cells (pDC)"
author: "Kim Dill-McFarland, kadm@uw.edu"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE}
#Create directories of outputs
dir.create("publication/fig/", showWarnings = FALSE, recursive = TRUE)
dir.create("publication/table/", showWarnings = FALSE, recursive = TRUE)

#Load packages used across all data
library(tidyverse)
library(limma)
#Set seed
set.seed(4389)
```

# Tables
## RNA-seq library and patient metadata

```{r}
attach("data_clean/P259_pDC_clean.RData")

dat.pDC.voom$targets %>% 
  select(libID, everything()) %>% 
  arrange(libID) %>% 
  write_csv(., "publication/table/TableS.metadata.csv")
```

### Patient summary demographics

```{r}
#Age
dat.pDC.voom$targets %>% 
  distinct(experiment, donorID, Age) %>% 
  group_by(experiment) %>% 
  summarise(.groups = "keep",
            n = n(),
            mean.age = mean(Age, na.rm=TRUE),
            sd.age = sd(Age, na.rm=TRUE)/sqrt(n))
#Sex
dat.pDC.voom$targets %>% 
  distinct(experiment, donorID, Sex) %>% 
  count(experiment, Sex) %>% 
  group_by(experiment) %>% 
  mutate(pct = n/sum(n))
```

## RNA-seq normalized log2 counts

```{r}
#attach("data_clean/P259_pDC_clean.RData")

as.data.frame(dat.pDC.voom$E) %>% 
  rownames_to_column("geneName") %>% 
#Save
write_csv(., "publication/table/TableS.norm.log2.counts.csv")
```

## All contrast model results

```{r message=FALSE}
#Get all model results
read_csv("results/gene_level/P259.1_gene_pval.csv") %>% 
  #Add experiment variable
  mutate(experiment = "P259_1") %>% 
  bind_rows(read_csv("results/gene_level/P259.2_gene_pval.csv")) %>% 
  #Fill in experiment variable
  mutate(experiment = ifelse(is.na(experiment), "P259_2", experiment)) %>% 
  #Keep only contrasts model 
  filter(model == "contrasts") %>% 
  #Reorder and rename variables
  select(experiment, model, geneName, hgnc_symbol, group,
         AveExpr:group) %>% 
  rename(contrast = group) %>% 
  arrange(experiment, geneName, contrast) %>% 
#Save
write_csv(., "publication/table/TableS.models.csv")
```

## GSEA results

```{r message=FALSE}
read_csv("results/GSEA_FoldChange/h_GSEA.result.csv") %>% 
  rename(contrast = group) %>% 
  #Add experiment variable
  mutate(experiment = ifelse(grepl(".1", contrast), "P259_1", "P259_2")) %>%
  #Set gene set variable
  mutate(gene.set = "HALLMARK",
         pathway = gsub("HALLMARK_", "", pathway)) %>% 
  #Format contrast to match model results
  mutate(contrast = gsub(".1", "", contrast),
         contrast = gsub(".2", "", contrast),
         contrast = gsub("[.]n"," - n", contrast),
         contrast = gsub("[.]A"," - A", contrast),
         contrast = gsub("[.]E"," - E", contrast)) %>% 
  #Reorder variables
  select(experiment, gene.set, pathway, contrast, everything()) %>% 
  arrange(experiment, pathway, contrast) %>% 
#Save
write_csv(., "publication/table/TableS.GSEA.csv") 
```

# Figures
## PCA

```{r message=FALSE}
library(cowplot)
```

Calculate PCA for P259.1 and P259.2

```{r}
#attach("data_clean/P259_pDC_clean.RData")

PCA1 <- as.data.frame(dat.pDC.voom_1$E) %>% 
  t() %>% 
  prcomp(scale. = TRUE)

#Extract axes labels
PC1.label1 <- paste("PC1 (", 
                    round(summary(PCA1)$importance[2,1]*100, digits=1), 
                    "%)", sep="")
PC2.label1 <-paste("PC2 (", 
                   round(summary(PCA1)$importance[2,2]*100, digits=1), 
                   "%)", sep="")

PCA2 <- as.data.frame(dat.pDC.voom_2$E) %>% 
  t() %>% 
  prcomp(scale. = TRUE)

#Extract axes labels
PC1.label2 <- paste("PC1 (", 
                    round(summary(PCA2)$importance[2,1]*100, digits=1), 
                    "%)", sep="")
PC2.label2 <-paste("PC2 (", 
                   round(summary(PCA2)$importance[2,2]*100, digits=1), 
                   "%)", sep="")
```

Combine experiments and metadata.

```{r}
# Extract PC values
PCA.dat <- as.data.frame(PCA1$x) %>% 
  rownames_to_column("libID") %>%
  bind_rows(rownames_to_column(as.data.frame(PCA2$x),"libID")) %>% 
  # Select PCs
  dplyr::select(libID, PC1, PC2) %>% 
  # Merge with metadata
  left_join(dat.pDC.voom$targets, by="libID")
```

Plot PCAs

```{r fig.width=8.5}
#Define ggplot colors
group.cols <- c("none:none"="#dadaeb",
                "none:AntiIL5"="#9e9ac8",
                "none:EOS.supp"="#54278f",
                "HRV:none"="#c7e9c0",
                "HRV:AntiIL5"="#74c476",
                "HRV:EOS.supp"="#006d2c",
                "flu:none"="#fdae6b",
                "flu:AntiIL5"="#e6550d")

PCA1 <- PCA.dat %>% 
  filter(experiment=="P259_1") %>%

  ggplot(aes(PC1, PC2, color=virus:IL5)) +
  geom_point(size=5) +
  #Beautify
  theme_classic(base_size = 16) +
  labs(x=PC1.label1, y=PC2.label1) +
  coord_fixed(ratio=1)  +
  scale_color_manual(values = group.cols, 
                     labels = c("media", "+ EOS sup", 
                              "+ RV", "+ RV + EOS sup"),
                     name="") +
  theme(legend.position = "bottom") +
  guides(color=guide_legend(nrow=2))

PCA2 <- PCA.dat %>% 
  filter(experiment=="P259_2") %>%

  ggplot(aes(PC1, PC2, color=virus:IL5)) +
  geom_point(size=5) +
  #Beautify
  theme_classic(base_size = 16) +
  labs(x=PC1.label2, y=PC2.label2) +
  coord_fixed(ratio=1) +
  scale_color_manual(values = group.cols, 
                     labels = c("media", "+ Anti-IL5", 
                                "+ RV", "+ RV + Anti-IL5"),
                     name="") +
  theme(legend.position = "bottom") +
  guides(color=guide_legend(nrow=2))

plot_grid(PCA1,PCA2, align = "hv", labels = c("A","B"))

#Save
ggsave("publication/fig/PCA.pdf", 
       plot_grid(PCA1,PCA2, align = "hv", labels = c("A","B")),
       width=10, height=5)
```

## Contrast model venn

```{r}
library(venn)
```

Genes significant for contrasts at FDR < 0.2

```{r message=FALSE, fig.width=8.5}
#Load pvalue results
pval_1 <- read_csv("results/gene_level/P259.1_gene_pval.csv") %>% 
  filter(model == "contrasts")
pval_2 <- read_csv("results/gene_level/P259.2_gene_pval.csv") %>% 
  filter(model == "contrasts")

#Set FDR cutoff
fdr.list <- c(0.2)

#Plot venns
for(fdr.cutoff in fdr.list){
  #P259.1
  venn.list_1 <- list()
  ##Signif for virus
  venn.list_1[["RV-infected\nvs media"]] <- pval_1 %>% 
    filter(group %in% c("none_HRV - none_none",
                        "EOS.supp_HRV - EOS.supp_none") & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  ##Signif for treatment
  venn.list_1[["- RV\nEOS supernatant\nvs none"]] <- pval_1 %>% 
    filter(group == "EOS.supp_none - none_none" & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  venn.list_1[["+ RV\nEOS supernatant\nvs none"]] <- pval_1 %>% 
    filter(group == "EOS.supp_HRV - none_HRV" & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  #P259.2
  venn.list_2 <- list()
  ##Signif for virus
  venn.list_2[["RV-infected\nvs media"]] <- pval_2 %>% 
    filter(group %in% c("none_HRV - none_none",
                        "AntiIL5_HRV - AntiIL5_none") & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  ##Signif for treatment
  venn.list_2[["- RV\nAnti-IL5 therapy\nvs none"]] <- pval_2 %>% 
    filter(group == "AntiIL5_none - none_none" & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  venn.list_2[["+ RV\nAnti-IL5 therapy\nvs none"]] <- pval_2 %>% 
    filter(group == "AntiIL5_HRV - none_HRV" & 
             adj.P.Val <= fdr.cutoff) %>% 
    distinct(geneName) %>% 
    select(geneName) %>% unlist(use.names = FALSE)

  #Save
  pdf(paste("publication/fig/contrast.venn_", fdr.cutoff, ".pdf", sep=""),
      height=5, width=12)
  par(mfrow = c(1,2)) 
  venn(ilab=FALSE, zcolor = "style",ilcs=1.5, sncs=1.5, box=FALSE,
       x=venn.list_1)
  venn(ilab=FALSE, zcolor = "style",ilcs=1.5, sncs=1.5, box=FALSE,
       x=venn.list_2)
  dev.off()

  #Print to Rmd
  par(mfrow = c(1,2)) 
  venn(ilab=FALSE, zcolor = "style",ilcs=1.5, sncs=1.5, box=FALSE,
       x=venn.list_1)
  venn(ilab=FALSE, zcolor = "style",ilcs=1.5, sncs=1.5, box=FALSE,
       x=venn.list_2)
}
```

## Hallmark GSEA 
### Enrichment score

```{r message=FALSE, fig.width=8.5}
#Load plotting function
source("scripts/GSEA.score.plot.R")

#Create plot
gsea.plot(gsea = read_csv("results/GSEA_FoldChange/h_GSEA.result.csv"),
          fdr.cut = 0.1, prefix = "H", print.plot = TRUE, 
          save.plot = TRUE, height = 5, width = 11, 
          outdir="publication/fig/")
```

### Mean fold change

```{r}
library(cowplot)
library(ggrepel)
```

```{r message=FALSE, warning=FALSE, fig.width=8.5, fig.height=10}
#List genes in each Hallmark term
myGO <- fgsea::gmtPathways("data_clean/Broad_gmt/h.all.v7.1.symbols.gmt")
GO.df <- plyr::ldply(myGO, rbind) %>% 
  rename(term = `.id`) %>% 
  pivot_longer(-term, values_to="hgnc_symbol")

#Expression data
#attach("data_clean/P259_pDC_clean.RData")

#### Format data ####
#Select genes of interest and format
## P259.2
h.combo.plot_2 <- as.data.frame(dat.pDC.voom_2$E) %>% 
  rownames_to_column("geneName") %>% 
  #Get HGNC symbols
  left_join(dat.pDC.voom_2$genes) %>% 
  select(hgnc_symbol, everything()) %>% 
  #Filter to genes in terms
  inner_join(GO.df) %>% 
  select(term, hgnc_symbol, geneName, everything(), 
         -c(`Previous symbols`:gene_biotype), -name) %>% 
  #long format
  pivot_longer(-c(term:geneName), names_to = "libID", 
               values_to = "expression") %>% 
  #Add metadata
  mutate(experiment = "P259.2") %>% 
  left_join(select(dat.pDC.voom_2$targets, donorID, libID, 
                   IL5, virus.detail))

## P259.1
h.combo.plot_1 <- as.data.frame(dat.pDC.voom_1$E) %>% 
  rownames_to_column("geneName") %>% 
  #Get HGNC symbols
  left_join(dat.pDC.voom_1$genes) %>% 
  select(hgnc_symbol, everything()) %>% 
  #Filter to genes in terms
  inner_join(GO.df) %>% 
  select(term, hgnc_symbol, geneName, everything(), 
         -c(`Previous symbols`:gene_biotype), -name) %>% 
  #long format
  pivot_longer(-c(term:geneName), names_to = "libID", 
               values_to = "expression") %>% 
  #Add metadata
  mutate(experiment = "P259.1") %>% 
  left_join(select(dat.pDC.voom_1$targets, donorID,
                   libID, IL5, virus.detail))

#Combine data and calculate fold change
h.combo.plot <- bind_rows(h.combo.plot_1, h.combo.plot_2) %>% 
  mutate(group = paste(virus.detail, IL5, sep=".")) %>% 
  select(-IL5, -virus.detail, -libID) %>% 
  #Averages
  group_by(term, hgnc_symbol, experiment, group) %>% 
  summarise(meanE = mean(expression, na.rm=TRUE)) %>% 
  #Fold change
  pivot_wider(names_from = group, values_from = meanE) %>% 
  rowwise() %>% 
  #Simplify group names
  mutate(
    #Virus in untreated
    virus.untreat_old = oldHRV.none-none.none,
    virus.untreat_new = newHRV.none-none.none,
    #Virus in treated
    virus.treat_old1 = oldHRV.EOS.supp-none.EOS.supp,
    virus.treat_old2 = oldHRV.AntiIL5-none.AntiIL5,
    virus.treat_new = newHRV.AntiIL5-none.AntiIL5,
    #Treatment in media
    treat.media_1 = none.EOS.supp-none.none,
    treat.media_2 = none.AntiIL5-none.none,
    #Treatment in virus
    treat.virus_old1 = oldHRV.EOS.supp-oldHRV.none,
    treat.virus_old2 = oldHRV.AntiIL5-oldHRV.none,
    treat.virus_new = newHRV.AntiIL5-newHRV.none) %>% 
  ungroup() %>%

  #long format
  select(term:experiment, virus.untreat_old:treat.virus_new) %>% 
  pivot_longer(virus.untreat_old:treat.virus_new) %>% 
  drop_na(value) %>%

  #Average old and new HRV values
  separate(name, into=c("group"), sep="_") %>% 
  group_by(term, hgnc_symbol, experiment, group) %>% 
  summarise(value = mean(value, na.rm=TRUE)) %>% 
  ungroup() %>% 
  #Beautify for plotting
  mutate(term = gsub("HALLMARK_","", term),
         term = gsub("_"," ", term)) %>% 
  mutate(col.group = ifelse(value <0, "down", "up")) %>% 
  mutate(group = factor(group, levels=c("treat.media","treat.virus",
                                        "virus.untreat","virus.treat")))

#Set jitter
pos <- position_jitter(width = 0.3, seed = 589, height = 0)

#### List genes to label ####
#Set max number of terms
no.genes <- 5
#Set gene sets of interest
term.OI.ls <- c("INTERFERON ALPHA RESPONSE", "INTERFERON GAMMA RESPONSE",
             "INFLAMMATORY RESPONSE",
             "EPITHELIAL MESENCHYMAL TRANSITION")

#Find genes with largest absolute fold change  
to.label <- list()

for(term.OI in term.OI.ls){
  #If this term, find genes that go down with virus
  if(term.OI == "EPITHELIAL MESENCHYMAL TRANSITION"){
    genes.v <- h.combo.plot %>% 
      filter(term ==term.OI &
               group %in% c("virus.treat","virus.untreat")) %>% 
      group_by(term, hgnc_symbol) %>% 
      filter(max(value, na.rm=TRUE)<0) %>% 
      summarise(max.fc = max(abs(value), na.rm=TRUE)) %>% 
      ungroup() %>% 
      slice_max(max.fc,n=no.genes) %>% 
      distinct(hgnc_symbol)
  } else{
    #All other terms, find genes that go up with virus
    genes.v <- h.combo.plot %>% 
      filter(term ==term.OI &
               group %in% c("virus.treat","virus.untreat")) %>% 
      group_by(term, hgnc_symbol) %>% 
      filter(min(value, na.rm=TRUE)>0) %>% 
      summarise(max.fc = max(value, na.rm=TRUE)) %>% 
      ungroup() %>% 
      slice_max(max.fc,n=no.genes) %>% 
      distinct(hgnc_symbol)
  }

  #Save genes to list object
  to.label[[paste("v",term.OI,sep=".")]] <- genes.v

  #Find genes that go down with EOS supp
  down.eos <- h.combo.plot %>% 
    filter(term == term.OI & group %in% c("treat.media","treat.virus") &
             experiment == "P259.1") %>% 
    group_by(term, hgnc_symbol) %>% 
    filter(max(value, na.rm=TRUE)<0) %>% 
    ungroup() %>% 
    distinct(hgnc_symbol) %>% unlist(use.names = FALSE)

  #Find genes that go up with AntiIL5
  up.aIL5 <- h.combo.plot %>% 
    filter(term == term.OI & group %in% c("treat.media","treat.virus") &
             experiment == "P259.2") %>% 
    group_by(term, hgnc_symbol) %>% 
    filter(min(value, na.rm=TRUE)>0) %>% 
    ungroup() %>% 
    distinct(hgnc_symbol) %>% unlist(use.names = FALSE)

  #Concatenate to genes that go up/down with virus and down with EOS and 
  # up with AntiIL5
  ## i.e. follow the directions in the enrichment score plot
  genes.t <- h.combo.plot %>% 
    filter(term == term.OI &
             group %in% c("virus.treat","virus.untreat")) %>% 
    filter(hgnc_symbol %in% intersect(down.eos,up.aIL5)) %>%

    group_by(term, hgnc_symbol) %>% 
    filter(max(abs(value), na.rm=TRUE)>0) %>% 
    summarise(max.fc = max(abs(value), na.rm=TRUE)) %>% 
    ungroup() %>% 
    slice_max(max.fc,n=no.genes) %>% 
    distinct(hgnc_symbol)

  #Save to list object
  to.label[[paste("t",term.OI,sep=".")]] <- genes.t
}

#Convert genes to label to df
to.label.df <- data.table::rbindlist(to.label, idcol="name") %>% 
  separate(name, into=c("vt","term"), sep="[.]") %>% 
  mutate(group1 = ifelse(vt == "t", "treat.media", "virus.untreat"),
         group2 = ifelse(vt == "t", "treat.virus", "virus.treat")) %>% 
  pivot_longer(group1:group2, values_to="group") %>% 
  dplyr::select(-name, -vt) %>% 
  mutate(to.label = "y")

#add genes to label info to other  plot data
h.combo.plot.lab <- h.combo.plot %>% 
  full_join(to.label.df, by = c("term", "hgnc_symbol", "group"))

### FC plot data ####
FC.plot.dat <- h.combo.plot.lab %>% 
  #Beautify labels
  mutate(group = paste(experiment, group, sep="_")) %>% 
  mutate(group1 = recode_factor(factor(group),
                  "P259.2_virus.treat"="+ Anti-IL5 therapy",
                  "P259.2_virus.untreat"="- Anti-IL5 therapy",
                  "P259.1_virus.treat"="+ EOS supernatant",
                  "P259.1_virus.untreat"="- EOS supernatant",

                  "P259.1_treat.virus"="+ RV",
                  "P259.2_treat.virus"="+ RV",
                  "P259.1_treat.media"="- RV",
                  "P259.2_treat.media"="- RV"),
         group2 = recode_factor(factor(group),
                  "P259.1_virus.untreat"="RV-infected vs media",
                  "P259.2_virus.untreat"="RV-infected vs media",
                  "P259.1_virus.treat"="RV-infected vs media",
                  "P259.2_virus.treat"="RV-infected vs media",

                  "P259.1_treat.virus"="EOS supernatant vs none",
                  "P259.2_treat.virus"="Anti-IL5 therapy vs none",
                  "P259.1_treat.media"="EOS supernatant vs none",
                  "P259.2_treat.media"="Anti-IL5 therapy vs none")) %>% 
    #Reorder terms
    mutate(term.ord = recode_factor(factor(term), 
  "INTERFERON ALPHA RESPONSE"="INTERFERON\nALPHA\nRESPONSE",
  "INTERFERON GAMMA RESPONSE"="INTERFERON\nGAMMA\nRESPONSE",
  "INFLAMMATORY RESPONSE"="INFLAMMATORY\nRESPONSE",
  "EPITHELIAL MESENCHYMAL TRANSITION"="EPITHELIAL\nMESENCHYMAL\nTRANSITION"))

#### FC plot: VIRUS ####
dat.v <- FC.plot.dat %>% 
  #Reorder terms
      filter(group2 == "RV-infected vs media" &
             term.ord %in% c("INTERFERON\nALPHA\nRESPONSE",
                           "INTERFERON\nGAMMA\nRESPONSE",
                           "INFLAMMATORY\nRESPONSE",
                           "EPITHELIAL\nMESENCHYMAL\nTRANSITION"))

plot.v <- dat.v %>% 
      ggplot(aes(x = group1, y=value)) +
      geom_violin() +
      #Add non-labeled points
      geom_jitter(data=filter(dat.v, is.na(to.label)),
                  aes(color=col.group), position = pos) +
      #Add labeled points
      geom_point(data=filter(dat.v, !is.na(to.label)),
                 aes(color=col.group), ) +
      #Add labels left
      geom_text_repel(data=filter(dat.v, !is.na(to.label)),
                      aes(label=hgnc_symbol), direction="y",
                      hjust=1, nudge_x=-0.4,
                      show.legend = FALSE, size=3) +
      #Add line at 0
      #geom_hline(yintercept = 0, size=0.5) +
      #Add mean lines
      stat_summary(fun="mean", geom="crossbar")+
      facet_grid(term.ord~group2, scales="free") +
      coord_flip() +
      #Beautify
      theme_bw() +
      labs(x="", y="Gene mean log2 fold change") +
      scale_color_manual(values=c("down"="#6baed6","up"="#ef3b2c")) +
      theme(strip.text.y = element_blank())+
      theme(legend.position = "none",
            panel.grid.major.y = element_blank(),
            panel.grid.minor.y = element_blank(),
            strip.background =element_rect(fill="white"))
    #plot.v

#### FC plot: TREATMENT ####
dat.t <- FC.plot.dat %>% 
  filter(group2 != "RV-infected vs media" &
           term.ord %in% c("INTERFERON\nALPHA\nRESPONSE",
                           "INTERFERON\nGAMMA\nRESPONSE",
                           "INFLAMMATORY\nRESPONSE",
                           "EPITHELIAL\nMESENCHYMAL\nTRANSITION"))

plot.t <- dat.t %>% 
  ggplot(aes(x = group1, y=value)) +
  geom_violin() +
  #Add non-labeled points
  geom_jitter(data=filter(dat.t, is.na(to.label)),
              aes(color=col.group), position = pos) +
  #Add labeled points
  geom_point(data=filter(dat.t, !is.na(to.label)),
             aes(color=col.group), ) +
  #Add labels left
  geom_text_repel(data=filter(dat.t, !is.na(to.label)),
                  aes(label=hgnc_symbol), direction="y",
                  hjust=1, nudge_x=-0.4,
                  show.legend = FALSE, size=3) +
  #Add line at 0
  #geom_hline(yintercept = 0, size=0.5) +
  #Add mean lines
  stat_summary(fun="mean", geom="crossbar")+
  facet_grid(term.ord~group2, scales="free") +
  coord_flip() +
  #Beautify
  theme_bw() +
  labs(x="", y="Gene mean log2 fold change") +
  scale_color_manual(values=c("down"="#6baed6","up"="#ef3b2c")) +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background =element_rect(fill="white"))
#plot.t

#### Combine FC plots ####
fc.plot.all <- plot_grid(plot.v, plot.t, ncol=2, rel_widths = c(0.6,1))
fc.plot.all
ggsave("publication/fig/H_GSEA_FoldChange.pdf", fc.plot.all, 
       width = 20, height = 15)
```

## Genes of interest

```{r}
# List all genes to plot
GOI <- c("CXCL10", "CXCL11")

#Get results for genes of interest
dat.GOI <- bind_rows(h.combo.plot_1, h.combo.plot_2) %>% 
  filter(hgnc_symbol %in% GOI) %>% 
  dplyr::select(-term) %>% distinct()

plot.GOI <- dat.GOI %>% 
  #Beautify labels
  mutate(x.lab=paste(virus.detail,IL5, sep="_"),
         x.lab=gsub("oldH|newH", "", x.lab),
         x.lab=recode_factor(factor(x.lab),
                             "none_none"="media",
                             "none_EOS.supp"="+ EOS sup",
                             "none_AntiIL5"="+ Anti-IL5",
                             "RV_none"="+ RV",
                             "RV_EOS.supp"="+ RV\n+ EOS sup",
                             "RV_AntiIL5"="+ RV\n+ Anti-IL5")) %>% 
  mutate(ex.lab = recode_factor(factor(experiment),
                      "P259.1"="EOS supernatant\nin healthy donors",
                      "P259.2"="Anti-IL5 therapy\nin asthma donors")) %>% 
  arrange(x.lab) %>%

  #Plot
  ggplot(aes(x=x.lab, y=expression, color=donorID)) +
  geom_jitter(width=0.1, height=0) +
  #Add mean and error bars
  stat_summary(fun.data=mean_sdl, 
               fun.args = list(mult=1), 
               geom="errorbar", color="black", width=0.1) +
  stat_summary(fun=mean, geom="errorbar", 
               aes(ymax=..y.., ymin=..y..),
               color="black", width=0.25) +
  #facet experiments and genes
  facet_grid(hgnc_symbol~ex.lab, scales="free") +
  #Beautify
  theme_bw() +
  labs(x="", y="Normalized log2 expression") +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background =element_rect(fill="white"))

plot.GOI
ggsave("publication/fig/GOI.pdf", plot.GOI,
       height=6, width=6)
```

# R session

```{r}
sessionInfo()
```

***