### 2022-03-18 3.P259.pDC_response.to.review.Rmd
Add response to JACI review and update figures

---
title: 'P259: Response to reviews'
subtitle: "Dendritic cells (pDC)"
author: "Kim Dill-McFarland, kadm@uw.edu"
date: "version `r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
editor_options:
  chunk_output_type: console
---
# Background

Analyses performed in response to reviews for publication in JACI.

# Setup
Load packages

```{r message=FALSE, warning=FALSE}
# Data manipulation and figures
library(tidyverse)
library(ggpubr)
library(readxl)
library(patchwork)
#Print pretty tables to Rmd
library(knitr)
library(kableExtra)
```

Set seed

```{r}
set.seed(589)
```

Custom functions

```{r}
#To extract and format p-value from lmFit
source("https://raw.githubusercontent.com/kdillmcfarland/R_bioinformatic_scripts/master/limma.extract.pval.R")
```

# Low-quality libraries

Determine if low-quality libraries removed from analysis are associated with low input RNA. All filtered libraries with low raw sequences and high CV coverage also had low input RNA.

```{r echo=FALSE, message=FALSE, fig.width=8.5}
#### Data ####
attach("data_clean/P259_pDC_clean.RData")
meta <- read_csv("data_clean/P259_pDC_metadata.csv")
anno <- read_excel("data_raw/P259-2 Final Annotation.xlsx")

dat <- meta %>% 
  select(libID, total_sequences, median_cv_coverage) %>% 
  inner_join(select(anno, "library sampleId", "RNA conc (ng/ul)"), 
             by=c("libID"="library sampleId")) %>% 
  mutate(`Quality filter` = ifelse(libID %in% dat.pDC.voom$targets$libID, "Pass", "Fail"))

#### Plot ####
plot1 <- dat %>% 
  ggplot(aes(y=`RNA conc (ng/ul)`, x=total_sequences)) +
  geom_point(aes(color=`Quality filter`), size=2) +
  theme_classic() +
  labs(x = "Raw sequences", y="RNA concentration (ng/ul)")

plot2 <- dat %>% 
  ggplot(aes(y=`RNA conc (ng/ul)`, x=median_cv_coverage)) +
  geom_point(aes(color=`Quality filter`), size=2) +
  theme_classic() +
  labs(x = "Raw sequences", y="RNA concentration (ng/ul)")

plot1+plot2
```

# Viral load
### Data

RV sequences were extracted from RNA-seq libraries and quantified using methods similar to human data. Normalized RV counts were calculated as total RV sequences / total non-human sequences * 1E6

```{r echo=FALSE, message=FALSE}
dat2 <- read_csv("data_raw/211102_P259-1_P259-2_RhinoViruses_normlized_read_counts.csv") %>% 
  rename(libID=libid) %>% 
  #filter libraries in final analysis
  filter(libID %in% dat.pDC.voom$targets$libID) %>% 
  #add metadata
  left_join(select(meta, libID, experiment, donorID, IL5, virus, virus.detail)) %>% 
  mutate(contrast=paste(IL5,virus, sep="_")) %>% 
  mutate(contrast=factor(contrast, 
                         levels = c("none_none","none_HRV",
                                 "AntiIL5_none","AntiIL5_HRV",
                                 "EOS.supp_none","EOS.supp_HRV")))
```

```{r echo=FALSE, message=FALSE}
# Format for limma
count1 <- dat2 %>% 
  filter(experiment == "P259_1") %>% 
  select(libID, RhinoVirusA_normCount) %>% 
  pivot_longer(-libID) %>% 
  arrange(libID) %>% 
  pivot_wider(names_from = libID) %>% 
  column_to_rownames("name")

meta1 <- dat2 %>% 
  filter(experiment == "P259_1") %>% 
  arrange(libID) %>% 
  droplevels()

count2 <- dat2 %>% 
  filter(experiment == "P259_2") %>% 
  select(libID, RhinoVirusA_normCount) %>% 
  pivot_longer(-libID) %>% 
  arrange(libID) %>% 
  pivot_wider(names_from = libID) %>% 
  column_to_rownames("name")

meta2 <- dat2 %>% 
  filter(experiment == "P259_2") %>% 
  arrange(libID) %>% 
  droplevels()
```

### Run limma contrasts model

In both experiments, viral loads were significantly higher in virus-infected samples and there were no differences between EOS supernatant or AntiIL5 treatment groups.

```{r echo=FALSE}
# Define model
model_1.contrast<- model.matrix(~ 0 + contrast, data=meta1)
    colnames(model_1.contrast) <- c(
                                 "none_none","none_HRV",
                                 "EOS.supp_none","EOS.supp_HRV")
#Block by donor  
consensus.corr1 <- duplicateCorrelation(
  count1,
  model_1.contrast,
  block=meta1$donorID)$consensus.correlation

#Fit model
fitQW_1.contrast <- lmFit(count1, 
                      model_1.contrast,
                      block=meta1$donorID,
                      correlation=consensus.corr1)

#Get contrasts
contrast.matrix1 <- makeContrasts(
  none_HRV-none_none, 
  EOS.supp_HRV-EOS.supp_none,
  EOS.supp_none-none_none,
  EOS.supp_HRV-none_HRV,
  levels=model_1.contrast)

efitQW_1.contrast <- eBayes(contrasts.fit(fitQW_1.contrast,
                                  contrast.matrix1))

extract.pval(model=model_1.contrast,
             voom.dat=count1, 
             eFit=efitQW_1.contrast, 
             name="pval_1.contrast",
             summary=FALSE,
             contrasts=TRUE,
             contrast.mat=contrast.matrix1)
```

```{r echo=FALSE}
# Define model
model_2.contrast<- model.matrix(~ 0 + contrast, data=meta2)
    colnames(model_2.contrast) <- c(
                                 "none_none","none_HRV",
                                 "AntiIL5_none","AntiIL5_HRV")
#Block by donor  
consensus.corr2 <- duplicateCorrelation(
  count2,
  model_2.contrast,
  block=meta2$donorID)$consensus.correlation

#Fit model
fitQW_2.contrast <- lmFit(count2, 
                      model_2.contrast,
                      block=meta2$donorID,
                      correlation=consensus.corr2)

#Get contrasts
contrast.matrix2 <- makeContrasts(
  none_HRV-none_none, 
  AntiIL5_HRV-AntiIL5_none,
  AntiIL5_none-none_none,
  AntiIL5_HRV-none_HRV,
  levels=model_2.contrast)

efitQW_2.contrast <- eBayes(contrasts.fit(fitQW_2.contrast,
                                  contrast.matrix2))

extract.pval(model=model_2.contrast,
             voom.dat=count2, 
             eFit=efitQW_2.contrast, 
             name="pval_2.contrast",
             summary=FALSE,
             contrasts=TRUE,
             contrast.mat=contrast.matrix2)
```

```{r echo=FALSE}
pval_1.contrast %>% 
  select(group, logFC, adj.P.Val) %>% 
  kable(align=c("l","c","c"), digits=4, caption="P259_1",
        col.names = c("Variable","Fold change", "FDR")) %>% 
  kable_styling(latex_options="HOLD_position", full_width = FALSE)

pval_2.contrast %>% 
  select(group, logFC, adj.P.Val) %>% 
  kable(align=c("l","c","c"), digits=4, caption="P259_2", 
        col.names = c("Variable","Fold change", "FDR")) %>% 
  kable_styling(latex_options="HOLD_position", full_width = FALSE)
```

```{r echo=FALSE}
#save
write_csv(pval_1.contrast, file="review_response/RV.EOS.model.csv")
write_csv(pval_2.contrast, file="review_response/RV.antiIL5.model.csv")
```

### Plot

```{r echo=FALSE, fig.width=8.5, message=FALSE, warning=FALSE}
#### Format pval data ####
#Pvals for eos experiment
GOI.p1 <- pval_1.contrast %>% 
  mutate(dataset="P259.1")

#Pvals for aniti-IL5 experiment
GOI.p <- pval_2.contrast%>% 
  mutate(dataset="P259.2") %>% 
  #Combine with eos data
  bind_rows(GOI.p1) %>% 
  #Make symbols for plots
  mutate(symbol = ifelse(P.Value <= 0.001,"***",
                         ifelse(P.Value <= 0.01, "**",
                                ifelse(P.Value <= 0.05, "*", NA)))) %>% 
  filter(!is.na(symbol)) %>% 
  #Match group labels to those in plots
  mutate(group.long = paste(dataset, group, sep="_"),
         group1 = recode_factor(factor(group.long),
                        "P259.2_none_HRV - none_none"='"-\n-"',
                        "P259.2_AntiIL5_HRV - AntiIL5_none"='"-\n+"',
                        "P259.2_AntiIL5_none - none_none"='"-\n-"',
                        "P259.2_AntiIL5_HRV - none_HRV"='"+\n-"',
                        "P259.1_none_HRV - none_none"='"RV            -\nEOS sup  -"',
                        "P259.1_EOS.supp_HRV - EOS.supp_none"='"-\n+"',
                        "P259.1_EOS.supp_none - none_none"='"RV            -\nEOS sup  -"',
                        "P259.1_EOS.supp_HRV - none_HRV"='"+\n-"'),
         group2 = recode_factor(factor(group.long),
                          "P259.2_none_HRV - none_none"='"+\n-"',
                          "P259.2_AntiIL5_HRV - AntiIL5_none"='"+  RV\n+  Anti-IL-5/5R"*alpha',
                          "P259.2_AntiIL5_none - none_none"='"-\n+"',
                          "P259.2_AntiIL5_HRV - none_HRV"='"+  RV\n+  Anti-IL-5/5R"*alpha',
                          "P259.1_none_HRV - none_none"='"+\n-"',
                          "P259.1_EOS.supp_HRV - EOS.supp_none"='"+\n+"',
                          "P259.1_EOS.supp_none - none_none"='"-\n+"',
                          "P259.1_EOS.supp_HRV - none_HRV"='"+\n+"'),
         facet.lab = recode_factor(factor(dataset),
                            "P259.1" = '"EOS sup"',
                            "P259.2" = '"Anti-IL-5/5R"*alpha')) %>% 
  dplyr::select(facet.lab, group1, group2, symbol)

#Add y location for pval based on max expression in plot
GOI.pe <- dat2 %>% 
  #Max expression per gene and experiment
  group_by(experiment) %>% 
  summarise(max.e = max(RhinoVirusA_normCount, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate(facet.lab = experiment,
         facet.lab = recode_factor(factor(facet.lab),
                            "P259_1" = '"EOS sup"',
                            "P259_2" = '"Anti-IL-5/5R"*alpha')) %>% 
  #Add to pval data
  right_join(GOI.p) %>% 
  arrange(experiment, group1, group2)

#first entry per gene
#Set y position to 1
first <- GOI.pe %>% 
  group_by(experiment) %>% 
  slice(1) %>% 
  mutate(y.position1 = 1)

#Add first position data back and fill in remaining
GOI.pey <- GOI.pe %>% 
  full_join(first) %>%

  #Fill in positions 2 - N
  group_by(experiment, facet.lab) %>% 
  mutate(y.position2 = lag(y.position1)+1,
         y.position3 = lag(y.position2)+1,
         y.position4 = lag(y.position3)+1) %>% 
  #Collapse positions into 1 column
  mutate(y.position = ifelse(!is.na(y.position1),y.position1,
                             ifelse(!is.na(y.position2),y.position2,
                                    ifelse(!is.na(y.position3),y.position3,
                                           ifelse(!is.na(y.position4),y.position4,NA))))) %>% 
  #Scale to max expression
  mutate(y.position = max.e+y.position*5E3) %>% 
  ungroup()
#plot
plot3 <- dat2 %>% 
  mutate(x.lab=paste(experiment, virus.detail, IL5, sep="_"),
         x.lab=gsub("oldH|newH", "", x.lab),
         x.lab=recode_factor(factor(x.lab),
                             "P259_1_none_none"='"RV            -\nEOS sup  -"',
                             "P259_2_none_none"='"-\n-"',
                             "P259_1_none_EOS.supp"='"-\n+"',
                             "P259_2_none_AntiIL5"='"-\n+"',
                             "P259_1_RV_none"='"+\n-"',
                             "P259_2_RV_none"='"+\n-"',
                             "P259_1_RV_EOS.supp"='"+\n+"',
                             "P259_2_RV_AntiIL5"='"+  RV\n+  Anti-IL-5/5R"*alpha')) %>% 
  mutate(facet.lab = experiment,
         facet.lab = recode_factor(factor(facet.lab),
                            "P259_1" = '"EOS sup"',
                            "P259_2" = '"Anti-IL-5/5R"*alpha')) %>% 
  arrange(facet.lab) %>% 
  ggplot(aes(x=x.lab, y=RhinoVirusA_normCount, color=donorID)) +
  geom_jitter(width=0.1, height=0) +
  stat_summary(fun.data=mean_sdl, 
               fun.args = list(mult=1), 
               geom="errorbar", color="black", width=0.25) +
  stat_summary(fun=mean, geom="errorbar", 
               aes(ymax=..y.., ymin=..y..),
               color="black", width=0.5) +
  facet_grid(~facet.lab, scales="free", 
             labeller = labeller(facet.lab=label_parsed)) +
  #Add pval
  stat_pvalue_manual(data=GOI.pey,
                     label="symbol", xmin="group1", xmax="group2") +
  # #Beautify
  theme_bw() +
  labs(x="", y="Normalized RV expression") +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background =element_rect(fill="white"),
        axis.text.x=element_text(hjust=0.95,vjust=-0.5),
        plot.margin = margin(0.2,0.2,0.2,0.2,"cm")) +
  scale_x_discrete(labels = ggplot2:::parse_safe)
plot3
```

# R session

```{r}
sessionInfo()
```

***